---
title: "OccupationalProfiles_Analysis"
author: "Barbara Hofer"
date: "31 7 2019"
output:
  pdf_document:
    df_print: paged
  html_document:
    df_print: paged
toc: TRUE    
urlcolor: blue
---

\newpage

### License
not yet specified

### Acknowledgements
This work was supported by the Sector Skill Alliance project EO4GEO - http://www.eo4geo.eu Many thanks to all EO4GEO partners who contributed occupational profiles.

### Metadata
Required libraries and runtime environment description.

```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library("here")
library("dplyr")
library("stringr")
library("DT")
library("data.table")
library("formattable")
library("htmltools")
library("devtools")
library("knitr")
```

```{r session_info, echo=FALSE, warning=FALSE}
# wrap long lines in R:
opts_chunk$set(tidy.opts=list(width.cutoff=70),tidy=TRUE)
```
\newpage

### Data
The data used in the analysis are the .csv file named `Profiles_inputdata_180719.csv`. This file has been generated based on occupational profiles that have been contributed by partners of the EO4GEO project. The file contains the duties and tasks with in some cases newly assigned labels of profiles related to Remote sensing and GIS workforce.

```{r load data, echo=TRUE}

#to set the path to the directory where the code resides
here::here("Profile_extendedanalysis_test", "Github_Material_July2019")

list.files(path='.')

profiles <- read.csv2("Profiles_inputdata_300719.csv", header = TRUE, sep =";")

#---
#coherence of strings in the duties and tasks
#this step is only done for the first manipulation of duties and tasks; manual adaptations have been done afterwards
#if statement checks, whether the colum of duty.man is empty and executes the code then
#---

if (sum(profiles$duty.man == "") == length(profiles$duty)) {

#for duties: to lower case, ignore everything in brackets, & replace by and, delete white spaces at the end of string
profiles$duty.man <- str_to_lower(profiles$duty, locale = "en")
profiles$duty.man <- str_replace(profiles$duty.man, "\\(.*\\)", "")
profiles$duty.man <- str_replace(profiles$duty.man, "&", "and")
profiles$duty.man <- str_trim(profiles$duty.man)

#for tasks: to lower case, ignore everything in brackets, & replace by and, delete white spaces at the end of string
profiles$task.man <- str_to_lower(profiles$original.task.name, locale = "en")
profiles$task.man <- str_replace(profiles$task.man, "\\(.*\\)", "")
profiles$task.man <- str_replace(profiles$task.man, "&", "and")
profiles$task.man <- str_trim(profiles$task.man)

#write.csv2(profiles, "Profiles_inputdata_manipulated15042019.csv")

}



```

```{r identify number of profiles, echo=FALSE}
nprofiles <- length(unique(profiles$profile.name))

```

``` {r comparison of original and manipulated duty columns}

#number of original duties and manipulated duties:
nduty <- length(unique(profiles$duty))
nduty.man <- length(unique(profiles$duty.man))

#number of (original) duties
nduty

#number of manipulated duties
nduty.man

#table summarizing the manipulation of the duties
dutynames <- profiles %>% group_by(profiles$duty, profiles$duty.man) %>% summarize()

datatable(dutynames)


```

``` {r comparison of original and manipulated task columns}

ntasks <- length(unique(profiles$original.task.name))
ntasks.man <- length(unique(profiles$task.man))

# number of (original) tasks
ntasks

# number of manipulated tasks  
ntasks.man

#generate a table for the original and manipulated tasks
tasknames <- profiles %>% group_by(  profiles$original.task.name, profiles$task.man) %>% summarize()
datatable(tasknames)

```

```{r calculate frequency of duties}

#I do a grouping by profile and duties and summarize the values
#the grouping is necessary because I have a duty entry for every task (row); so I summarize by original.profile to have a non-redundant list of duties
profduty <- profiles %>% group_by( profiles$duty.man, profiles$profile.name) %>% summarize()

#here I count the number of different duties in the created variable
nduty <- count((profduty[,1]))

#here the frequency of duties is calculated (in percent of total)
dutyfreq <- (nduty[,2]/nprofiles)*100

#add frequency to duties
nduty["dfreq"] <- dutyfreq[,1]

#inner join of profiles with nduty data frame (column names need to be identical)
nduty <- nduty %>% rename("duty.man" = "profiles$duty.man")
profilesd <- merge(profiles, nduty, by="duty.man")


#order the profiles according to the frequency of duties

oprofiles <- profilesd %>% group_by(profilesd$duty.man, profilesd$dfreq) %>% summarize()

#renaming of columns
oprofiles <- oprofiles %>% rename("duty.man" = "profilesd$duty.man")
oprofiles <- oprofiles %>% rename("dfreq" = "profilesd$dfreq")

#round percentage values
oprofiles$dfreq <-  round(oprofiles$dfreq,0) 

#ordering by dfreq
oprofiles <- oprofiles[order(oprofiles$dfreq, decreasing = T),]

```
### Frequency of duties over the profiles
The frequency of duties is provided in the following table: 

```{r result duty frequency }
datatable(oprofiles)


```
\newpage

### Cross section through the profiles per duty

In the following, two directories are created and filled with HTML table that include the tasks of single duties. The tasks are thereby either colored according to the indicate skill level that is required or according to trends that were identified. This results in two collections of 19 tables each that provide the main output of this analysis.


```{r generate single task collection for each duty, warning=FALSE}
# I build on the 19 duties and extract tasks for each profile in the original order
# I need the set of duties to iterate over
# in each iteration, I need to identify the profiles containing the duties and select the tasks of the duty
# then display the tasks according to the given order 
dir.create("Duties_skills")
dir.create("Duties_trends")

nduties <- length(unique(profiles$duty.man))

dutylist <- unique(profiles$duty.man)

dutylist <- str_replace(dutylist, "/", "or")
profiles$duty.man <- str_replace(profiles$duty.man, "/", "or")

menge <- data.frame()

remove(proftasks)
remove(proftasksp)
remove(proftasksskill)
remove(proftaskstrend)


for (d in 1:length(dutylist)) {
  proftasks <- data.frame()
  proftaskstrend <- data.frame()
  proftasksskill <- data.frame()
  
  menge <- profiles[profiles$duty.man == dutylist[d],]
  menge <- menge[order(menge$profile.source, menge$profile.name, menge$task.order),]
 
  sourceprof <- menge %>% group_by(menge$profile.source, menge$profile.name) %>% summarize()
  sourceprof <- sourceprof %>% rename("profile.source" = "menge$profile.source")
  sourceprof <- sourceprof %>% rename("profile.name" = "menge$profile.name")
  
  for (p in 1:nrow(sourceprof)) {
    
    taskset <- select(menge[menge$profile.name == sourceprof$profile.name[p],], "task.man")
    tasksettrend <- select(menge[menge$profile.name == sourceprof$profile.name[p],], "future.trend")
    tasksetskill <- select(menge[menge$profile.name == sourceprof$profile.name[p],], "skill.level..s.t.c.")
     
    taskdf <- data.frame(taskset)
    taskdf <- transpose(taskdf)
    
    taskdft <- data.frame(tasksettrend)
    taskdft <- transpose(taskdft)
    
    taskdfs <- data.frame(tasksetskill)
    taskdfs <- transpose(taskdfs)
    

    profilename <- merge(paste(sourceprof$profile.source[p]), paste(sourceprof$profile.name[p]))
    proftasksp <- merge(profilename, taskdf)
    proftasks <- rbindlist(list(proftasks, proftasksp), fill = TRUE)
    
    proftaskstrend <-  rbindlist(list(proftaskstrend, taskdft), fill = TRUE)
    
    proftasksskill<-  rbindlist(list(proftasksskill, taskdfs), fill = TRUE)
    
  }
  
  
  #write.csv2(proftasks,paste("Duties_skills/",dutylist[d], ".csv", sep=""), row.names = TRUE)
  
  

  #
  #Visualizing skills of tasks resulting in HTML tables
  #
  ntasks <- ncol(proftasks)
  nskills <- ncol(proftasksskill)
  
  colnames(proftasksskill) <- paste("SV", 1:nskills, sep = "") 
  
  taskvis <- cbind(proftasks, proftasksskill)
  
  taskvis[is.na(taskvis)] <- "-"
  

  n=(ncol(taskvis)-2)/2
  #this seems to work for the SV part of the list!
  SVcolumns <- do.call(list, lapply(1:n, function(i){
    return(FALSE)
  }))
  names(SVcolumns) <- paste("SV", 1:n, sep = "") 
  
  Vcolumns <- do.call(list, lapply(1:n, function(i){
    return(formatter("span", 
                     style = ~ style("background-color" = ifelse(taskvis[[paste("SV", i, sep="")]] =="s", "lemonchiffon", ifelse(taskvis[[paste("SV", i, sep="")]] =="c", "lightcoral", 
                                                                                                                                 ifelse(taskvis[[paste("SV", i, sep="")]] =="t", "mistyrose", 
                                                                                                                                        "white"))))))
    
  }))
  names(Vcolumns) <- paste("V", 1:n, sep = "") 
  
  SVandVformat <- c(SVcolumns, Vcolumns)
  
  #formattable(taskvis, align=rep("l", n), SVandVformat)
  
  
   #format_table provide the html version of the table (otherwise the function is called formattable)
   html_header="
  <head> 
  <meta charset=\"utf-8\"> 
  <meta name=\"viewport\" content=\"width=device-width, initial-scale=1\"> 
  <link rel=\"stylesheet\" href=\"https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css\">
  </head>
  <body>
  "
   html_table = format_table(taskvis, align=rep("l", n), SVandVformat)
   
   write(paste(html_header, html_table, sep=""), paste("Duties_skills/", dutylist[d], ".html", sep=""))
   
   #
   # Hier brauche ich noch ein paar oder um die Begriffe abzugrenzen und noch weitere Farben
   # und KOntrolle wichtig um zu sehen, ob die Trends richtig eingefärbt werden
   #
   #
   #Visualizing trends of tasks resulting in HTML tables
   #
   ntasks <- ncol(proftasks)
   ntrends <- ncol(proftaskstrend)
   
   colnames(proftaskstrend) <- paste("SV", 1:ntrends, sep = "") 
   
   taskvistrend <- cbind(proftasks, proftaskstrend)
   
   taskvistrend[is.na(taskvistrend)] <- "-"
   
   
   n=(ncol(taskvistrend)-2)/2
   #this seems to work for the SV part of the list!
   SVcolumns <- do.call(list, lapply(1:n, function(i){
     return(FALSE)
   }))
   names(SVcolumns) <- paste("SV", 1:n, sep = "") 
   
   Vcolumns <- do.call(list, lapply(1:n, function(i){
     return(formatter("span", 
                      style = ~ style("background-color" = ifelse(taskvistrend[[paste("SV", i, sep="")]] =="ias", "cadetblue", ifelse(taskvistrend[[paste("SV", i, sep="")]] =="pas", "lightsteelblue", 
                                                                                                                                  ifelse(taskvistrend[[paste("SV", i, sep="")]] =="analysis ready data", "lightyellow", 
                                                                                                                                         ifelse(taskvistrend[[paste("SV", i, sep="")]] =="automation of image analysis", "azure", 
                                                                                                                                                "white")))))))
     
   }))
   names(Vcolumns) <- paste("V", 1:n, sep = "") 
   
   SVandVformat <- c(SVcolumns, Vcolumns)
   
   formattable(taskvis, align=rep("l", n), SVandVformat)
   
   
   #format_table provide the html version of the table (otherwise the function is called formattable)
   html_header="
  <!DOCTYPE html>
  <html>
  <head> 
  <meta charset=\"utf-8\"> 
  <meta name=\"viewport\" content=\"width=device-width, initial-scale=1\"> 
  <link rel=\"stylesheet\" href=\"https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css\">
  </head>
  <body>
  "
   html_footer= "</html>"
   
   html_table = format_table(taskvistrend, align=rep("l", n), SVandVformat)
   
   write(paste(html_header, html_table, html_footer, sep=""), paste("Duties_trends/", dutylist[d], ".html", sep=""))
   
   
  remove(proftasks)
  remove(proftasksp)
}


```
